Inefficient Concessions and Mediation

### 2021-01-21 draft.tex
Revise Section 4.2; Thm 4 not quite finished, but proof moved

%Though it will not be modeled here, a stochastic component -- when countries occasionally unilaterally defect -- can explain why powerful countries like the USSR more often ``get away'' with cheating: the payoffs to the relationship are not symmetric. Losing the USA or USSR as an ally is much more devastating for a satellite country than vise versa, therefore big countries can have their alliances and take advantage of them too.
\subsection{Concessions Separating Equilibrium}
The other separating equilibrium of interest involves distrustful countries to using concessions as a way of building trust. We call this the \emph{concessions separating equilibrium}. In this equilibrium, the cost of giving a concession can be worth the investment if the other country also cooperates. It is also not as risky as the \emph{no-concession separating equilibrium}. The concessions are given before more substantive actions, thus concessions are a costly signal that the other country is serious about cooperating. Most agreements come with concessions, and in many cases, though not all, they are used as signals. The goal is to use this option to avoid continual distrust. That is, high types make use of concessions to identify themselves as trustworthy partners in hopes of establishing Peace.

@@

We implement this by allowing countries to choose a scalar $e$ to multiply by their given concession $g$. $0 \leq e \leq 1$. As before, $C(g)= g$ but while the receiving country observes the full concession $g$, its benefit is now only $eg$.\footnote{In a more general model, we could also think of $e$ as a function $E(\cdot)$. The argument is still $g$, but instead of a scalar, $E(\cdot)$ can transform $g$ into any concession, not only a numerical value. These gifts will be converted into utilities by cost and benefit functions, which then allow the possibility of a gift having efficiency greater than 1. The transformation function $E(\cdot)$ can be thought of as the type of gift, and has a cost to the sender and a direct benefit to the receiver in addition to its signaling value.}

@@

Before turning to the question of how the possibility of destroying the value of concessions impacts the ability of the parties to separate, we first formally establish the welfare implications.

\begin{lemma}
	Assuming that a separating equilibrium exists when concessions have material value and are efficient (where $e=1$), it is optimal to give efficient concessions in a separating equilibrium.
	\label{lemma:efficiency}
\end{lemma}
\emph{Proof:}
The benefit of a gift appears on both sides of the IC for both types and thus cancels out in the low type IC $(peg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+peg_h )$ and in the high type IC:  $(pX_{TT}^h+(1-p)X_{FF}^h - g_h +peg_h \geq X_{FF}^h +peg_h)$.\footnote{Elaboration and explanation of the IC constraints is in the proof of Theorem~\ref{theorem:2}.} Therefore, receiving a gift where $e<1$  does not affect incentives to truth tell about the countries' types. However, the expected separation utilities under this game are now
\be
U_h= pX_{TT}+(1-p)X_{FF} - g_h+peg_h 
\ee
and
\be
U_l=X_{FF}+peg_h 
\ee
 $\frac{dU}{de}>0$ for both countries so it is never optimal have $e<1$.\footnote{This result does not depend on the benefit so it holds for any monotonic benefit function. It also holds for any monotone cost function.} \hfill $\blacksquare$

@@

@@

\noindent {\bf \emph{Pooling equilibrium}}
\begin{lemma}
	From period 1 on, playing fight in all periods regardless of beliefs constitutes an equilibrium strategy of the continuation game for both types.
	\label{lemma:1}
\end{lemma}
\emph{Proof}: As $D\geq0$ and $W\geq0$ the dominant stage game action for all types is to play Fight. Thus the stage game equilibrium is (Fight, Fight) and in any repeated game, playing the stage game equilibrium in each period is an equilibrium of the entire game. \hfill $\blacksquare$\\

@@

That is, one high type equilibrium strategy is to pool with the low types---who always play ``Fight'' by definition---and always play Fight. If both the low and high type of player $j$ choose the strategy that gives no concessions and always plays fight, then either type of player $i$ is made strictly better off by playing Fight in each period given player $j$'s behavior. Given that the stage-game equilibrium will be (Fight,Fight), there is no incentive to give a concession in period 0 because concessions are costly and signaling one's type can give no benefit in this equilibrium. Types pool by never giving concessions and always playing Fight. 
 %add intro about the 3 types of equilibria. 
 \\
 \\
 {\bf \emph{No-Concessions Separating Equilibrium}}

@@

If the potential advantages of Peace are large enough to outweigh the risks of encountering a low type in expected utility terms, that is $\de_h \geq \frac{W}{(1-p)W + p(T+D)} = \de_h^{nc}$, a county has the incentive to play Trust in the first round rather than Fight. Any player with $\de < \de_h^{nc}$ has the incentive---by this same inequality---to play Fight in each period even if the other country plays Trust. Thus we call these countries low types for purposes of this equilibrium. \hfill $\blacksquare$\\
%\noindent{\emph{Proof of Corollary \ref{corollary:1}}:\\
%High-type welfare under the no-concessions separating equilibrium is $$\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D) \right\} = \frac{1}{1-\de_h}\left\{pT + (1-p)\left[W-D + \de_h(W-D)\right]\right\}$$
%High-type welfare under the grim trigger repeated game equilibrium is $\frac{1}{1-\de_h}\left\{pT + (1-p)\left[-D + \de_h(W-D) \right]\right\}$
%When $e=0$, welfare is $$\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$$
%This type of separating equilibrium can exist for some sets of parameters. Since the high types have sufficient incentive to separate under appropriate parameter values even without concessions being given in period 0, we refer to such an equilibrium as a \emph{no-concessions separating equilibrium}. This is why the threshold patience level for this type of equilibrium is labeled $\de_h^{nc}$.
\\
{\bf \emph{Concessions Separating Equilibrium}}

@@

In the case that both types of equilibria exist, i.e. $\de_h \geq \de_h^{nc}$, the high types will prefer the \emph{concessions separating equilibrium} if and only if
\begin{equation*}
p\left(\frac{T}{1- \delta_h}\right)+(1-p)\left(\frac{W-D}{1- \delta_h}\right) - (1-p)g_h \geq p\left(\frac{T}{1- \delta_h}\right) +(1-p)\left({-D}+\frac{\delta_h(W-D)}{1- \delta_h} \right)
\end{equation*}
This simplifies to preferring the concessions separating equilibrium if and only if
$$(1-p)W-p(1-p)(T+D)\geq 0$$
Rearranging,
\be
 \frac{W}{T+D}\geq p
\ee
i.e. The high types prefer the \emph{concessions separating equilibrium} when $p$ is smaller than $\frac{W}{T+D}$. \hfill $\blacksquare$\\
\\
{\bf \emph{Concessions Separating Equilibrium with Material Value}}\\
\emph{Proof of Theorem \ref{theorem:csemv}:}\\
In the separating equilibrium constructed here, players will believe that any concession not equal to the equilibrium concession of the high type $g^1$ is a low type gift. In this case, the binding incentive compatibility constraint is the low type IC constraint, which when simplified to allow $g_l=0$ following Lemma~\ref{lemma:3} is: 
\begin{equation}
	pg^1 +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g^1+pg^1
	\label{eq:matval1}
\end{equation}
The $\al_i$'s vary depending on the situation. On the left side, the low type does not give a concession but receives a concession if facing a high type ($pg^1$). Then (Fight, Fight) is played, and the payoffs depend on whether this low type is facing a high type or a low type because the implications for future material value are different. With probability $p$, the low type received a gift and invested in the military. With probability $(1-p)$ there were no concessions given by either side. The left side of Expression~\ref{eq:matval1} is then $pg^1 + \frac{1}{1-\de_l}\left[p(W+Wg^1-D) + (1-p)(W-D) \right]$.

@@

Substituting the minimum efficient concession from Expression~\ref{eq:matval2}, we have the condition for a concessions separating equilibrium to exist in terms of fundamentals:
\be
	\de_h \geq \frac{p(W-D-T)}{\frac{\left(1-\de_l\right)p\left(D +	 T\right)}{\left(1-\de_l\right)\left(1 - pT \right) + (1-p) D}} + p (W-T) + (1-p)D + 1
\ee
That is, there is no parameter restriction beyond our definition of what it means for a player to be a high type. \hfill $\blacksquare$\\
\\
{\bf \emph{Mediation}}\\
\emph{Proof of Theorem \ref{theorem:4}:}\\
High types choose $\alpha=1$ (because a concession would only come from another high type in equilibrium) and low types choose $\alpha=0$. Because the low type's IC constraint binds, with $e=1$ the optimal concession can be calculated to be $g^1= \frac{p(1-d_l)(D+T)}{(1-d_l)(1-pT)+(1-p)D}$. The high type's utility when $e=1$ is then $\frac{1}{1-d_h}\left\{pT(1+g^*) + (1-p)(W-D(1+g^*))\right\} - g^1$.

When $e=0$---that is, the future material value of the concessions is destroyed---we are essentially back in the case of Section~\ref{sec:bench} and the minimum separating concession is $g^0=p(T+D)$. The threshold for high types to cooperate is again $\frac{W}{T+D}$. High type utility when $e=0$ is $\frac{1}{1-d_h}\left\{pT + (1-p)(W-D)\right\} - g^0$.

@@

>>>>

%Though it will not be modeled here, a stochastic component -- when countries occasionally unilaterally defect -- can explain why powerful countries like the USSR more often ``get away'' with cheating: the payoffs to the relationship are not symmetric. Losing the USA or USSR as an ally is much more devastating for a satellite country than vise versa, therefore big countries can have their alliances and take advantage of them too.
\subsection{Concessions Separating Equilibrium}
\label{sec:cse}
The other separating equilibrium of interest involves distrustful countries to using concessions as a way of building trust. We call this the \emph{concessions separating equilibrium}. In this equilibrium, the cost of giving a concession can be worth the investment if the other country also cooperates. It is also not as risky as the \emph{no-concession separating equilibrium}. The concessions are given before more substantive actions, thus concessions are a costly signal that the other country is serious about cooperating. Most agreements come with concessions, and in many cases, though not all, they are used as signals. The goal is to use this option to avoid continual distrust. That is, high types make use of concessions to identify themselves as trustworthy partners in hopes of establishing Peace.

@@

We implement this by allowing countries to choose a scalar $e$ to multiply by their given concession $g$ where $0 \leq e \leq 1$. As before, $C(g)= g$ but while the receiving country observes the full concession $g$, the future benefit to the recipient is now only $eg$.\footnote{In a more general model, we could also think of $e$ as a function $E(\cdot)$. The argument is still $g$, but instead of a scalar, $E(\cdot)$ can transform $g$ into any concession. In particular, this allows the transformation to increase the value of the concession. The transformation function $E(\cdot)$ can be thought of as the type of gift, and has a cost to the sender and a direct benefit to the receiver in addition to its signaling value.} Table 3 provides the modified payoffs in the repeated game.

@@

Before turning to the question of how the possibility of destroying the value of concessions impacts the ability of the parties to achieve peace, we first formally establish the welfare implications.

\begin{lemma}
	If a separating equilibrium exists when concessions have material value and are efficient (where $e=1$), it is optimal to give efficient concessions in a separating equilibrium if and only if $pT \geq (1-p)D$.
	\label{lemma:efficiency}
\end{lemma}
\emph{Proof:} See Appendix.

@@

The intuition for Lemma~\ref{lemma:efficiency} is that future material value of concession helps you when your negotiating partner turns out to be a high type, and this happens $p$ of the time. On other other hand, the material value of your concession \emph{hurts} you when your negotiating partner turns out to be a low type. The future material of concessions increase the expected utility from a separating equilibrium when both the payoffs from Trust and the probability of a high type are relatively large.

@@

\subsection{Pooling equilibrium}
\begin{lemma}
	From period 1 on, playing fight in all periods regardless of beliefs constitutes an equilibrium strategy of the continuation game for both types.
	\label{lemma:1}
\end{lemma}
\emph{Proof}: As $D\geq0$ and $W\geq0$ the dominant stage game action for all types is to play Fight. Thus the stage game equilibrium is (Fight, Fight) and in any repeated game, playing the stage game equilibrium in each period is an equilibrium of the entire game. \hfill $\blacksquare$\\

@@

That is, one high type equilibrium strategy is to pool with the low types---who always play ``Fight'' by definition---and always play Fight. If both the low and high type of player $j$ choose the strategy that gives no concessions and always plays fight, then either type of player $i$ is made strictly better off by playing Fight in each period given player $j$'s behavior. Given that the stage-game equilibrium will be (Fight,Fight), there is no incentive to give a concession in period 0 because concessions are costly and signaling one's type can give no benefit in this equilibrium. Types pool by never giving concessions and always playing Fight. 
 %add intro about the 3 types of equilibria. 
 \\
 \\
\subsection{No-Concessions Separating Equilibrium}

@@

If the potential advantages of Peace are large enough to outweigh the risks of encountering a low type in expected utility terms, that is $\de_h \geq \frac{W}{(1-p)W + p(T+D)} = \de_h^{nc}$, a county has the incentive to play Trust in the first round rather than Fight. Any player with $\de < \de_h^{nc}$ has the incentive---by this same inequality---to play Fight in each period even if the other country plays Trust. Thus we call these countries low types for purposes of this equilibrium. \hfill $\blacksquare$
%\noindent{\emph{Proof of Corollary \ref{corollary:1}}:\\
%High-type welfare under the no-concessions separating equilibrium is $$\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D) \right\} = \frac{1}{1-\de_h}\left\{pT + (1-p)\left[W-D + \de_h(W-D)\right]\right\}$$
%High-type welfare under the grim trigger repeated game equilibrium is $\frac{1}{1-\de_h}\left\{pT + (1-p)\left[-D + \de_h(W-D) \right]\right\}$
%When $e=0$, welfare is $$\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$$
%This type of separating equilibrium can exist for some sets of parameters. Since the high types have sufficient incentive to separate under appropriate parameter values even without concessions being given in period 0, we refer to such an equilibrium as a \emph{no-concessions separating equilibrium}. This is why the threshold patience level for this type of equilibrium is labeled $\de_h^{nc}$.

\subsection{Concessions Separating Equilibrium}

@@

In the case that both types of equilibria exist, i.e. $\de_h \geq \de_h^{nc}$, the high types will prefer the \emph{concessions separating equilibrium} if and only if
\begin{equation*}
p\left(\frac{T}{1- \delta_h}\right)+(1-p)\left(\frac{W-D}{1- \delta_h}\right) - (1-p)g_h \geq p\left(\frac{T}{1- \delta_h}\right) +(1-p)\left({-D}+\frac{\delta_h(W-D)}{1- \delta_h} \right)
\end{equation*}
This simplifies to preferring the concessions separating equilibrium if and only if
$$(1-p)W-p(1-p)(T+D)\geq 0$$
Rearranging,
\be
 \frac{W}{T+D}\geq p
\ee
i.e. The high types prefer the \emph{concessions separating equilibrium} when $p$ is smaller than $\frac{W}{T+D}$. \hfill $\blacksquare$

\subsection{Concessions Separating Equilibrium with Material Value}
\emph{Proof of Theorem \ref{theorem:csemv}:}\\
In the separating equilibrium constructed here, players will believe that any concession not equal to the equilibrium concession of the high type $g^1$ is a low type gift. In this case, the binding incentive compatibility constraint is the low type IC constraint, which when simplified to allow $g_l=0$ following Lemma~\ref{lemma:3} is: 
\begin{equation}
	pg^1 +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g^1+pg^1
	\label{eq:matval1}
\end{equation}
The $\al_i$'s vary depending on the situation. On the left side, the low type does not give a concession but receives a concession if facing a high type ($pg^1$). Then (Fight, Fight) is played, and the payoffs depend on whether this low type is facing a high type or a low type because the implications for future material value are different. With probability $p$, the low type received a gift and invested in the military. With probability $(1-p)$ there were no concessions given by either side. The left side of Expression~\ref{eq:matval1} is then $pg^1 + \frac{1}{1-\de_l}\left[p(W+Wg^1-D) + (1-p)(W-D) \right]$.

@@

Substituting the minimum efficient concession from Expression~\ref{eq:matval2}, we have the condition for a concessions separating equilibrium to exist in terms of fundamentals:
\be
	\de_h \geq \frac{p(W-D-T)}{\frac{\left(1-\de_l\right)p\left(D +	 T\right)}{\left(1-\de_l\right)\left(1 - pT \right) + (1-p) D}} + p (W-T) + (1-p)D + 1
\ee
That is, there is no parameter restriction beyond our definition of what it means for a player to be a high type. \hfill $\blacksquare$

\subsection{Material Value of Concessions can be Destroyed}
\emph{Proof of Lemma~\ref{lemma:efficiency}:}\\
%The benefit of a gift appears on both sides of the incentive compatibility constraint for both types and thus cancels out in the low type incentive compatibility constraint $(peg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+peg_h )$ and in the high type incentive compatibility constraint: $(pX_{TT}^h+(1-p)X_{FF}^h - g_h +peg_h \geq X_{FF}^h +peg_h)$. Therefore, receiving a gift where $e<1$  does not affect incentives to truth tell about the countries' types.
The expected separation utility for the hight type under the game with material value that can be destroyed is
$$\frac{1}{1-\de_h}\left\{pT(1+eg^1) + (1-p)(W-D(1+eg^1))\right\} - (1-p)g^1$$
The derivative with respect to $e$ is
$$\frac{1}{1-\de_h}\left\{pTg^1 - (1-p)Dg^1)\right\}$$
This is non-negative when $pT \geq (1-p)D$. High type welfare is thus non-decreasing in $e$ when this condition is met.  The largest admissible value of $e$, i.e. $e=1$ thus maximizes welfare when $pT \geq (1-p)D$.\footnote{If $pT = (1-p)D$, $e=1$ is not the unique maximizer, but it is still a maximizer.} \hfill $\blacksquare$\\
%\footnote{This result does not depend on the benefit so it holds for any monotonic benefit function. It also holds for any monotone cost function.} 
\\
\emph{Proof of Theorem \ref{theorem:4}:}\\
High types choose $\alpha=1$ (because a concession would only come from another high type in equilibrium) and low types choose $\alpha=0$. The high type's utility when $e=1$ is
$$\frac{1}{1-\de_h}\left\{pT(1+g^1) + (1-p)(W-D(1+g^1))\right\} - (1-p)g^1 $$
where $g^1= \frac{p(1-d_l)(D+T)}{(1-d_l)(1-pT)+(1-p)D}$.

When $e=0$---that is, the future material value of the concessions is destroyed---we are essentially back in the case of Section~\ref{sec:cse}. High type utility when $e=0$ is $$\frac{1}{1-d_h}\left\{pT + (1-p)(W-D)\right\} - (1-p)g^0$$ where the minimum separating concession is $g^0=p(T+D)$. The threshold for high types to cooperate is again $\frac{W}{T+D}$.

@@

\subsection{Mediation}
